Leveraging Diversity to Find Bugs in JavaScript Engines

### 2019-02-20 paper/paper.tex
pass on evaluation

\emph{Key Findings.}~1) Not only multiple different implementations
can be leveraged in differential testing, but differences in test
suites can also be important. 2) Even for problems with fairly clear
specifications, as in \javascript{}, there is likely to be (a lot of)
variation between different implementations. 3) Differential testing
is feasible on real, complex, widely used pieces of software.

@@

\vspace{1ex} To sum, although more research needs be done to automate
parts of the inspection process (\eg{}, reduction of false alarms),
the paper provides initial, yet strong evidence that exploring
diversity should be encouraged to find functional bugs in JavaScript
engines.

@@

\section{JS Files Mined}
\label{sec:seeds}
A good test set is critical for finding bugs with the techniques used
in this paper. For that reason, we looked for \js\ files from various
sources: 1) test files from the Test262~\cite{tc39-github} conformance
suite of the ECMA262 specification~\cite{ecmas262-spec}, 2) test files
from the test suite of the selected engines (see
Section~\ref{sec:methodology:engines}); these files are accessible
from the engine's official repositories, 3) test files from the suites
of public engines, and 4) test files mined from issue trackers of
these engines.

@@

Figure~\ref{fig:workflow} illustrates the 
differential testing infrastructure used in part of this study. It takes on
input a list of JS files and generates warnings on output. Boxes in
the figure denote encapsulation, arrowed lines denote data flow, and
the cycle icons denote repetition--the leftmost icon indicates that
each file in the input list will be analyzed in separate whereas the
rightmost icon shows that a single file will be fuzzed multiple times.
The bug-finding process works as follows. New inputs are obtained for
a given test input using some off-the-shelf input
fuzzer; Section~\ref{sec:objects:fuzzers} describes the fuzzers we selected.
The oracle checks whether or not the output produced for the fuzzed
file is consistent across all engines. In case the test passes in all
engines or fails in all engines (\ie{}, the output is consistent), the
infrastructure ignores the input. Otherwise, it considers the input as
potentially fault-revealing; hence, interesting for human
inspection. The infrastructure uses the
open-source tool eshost-cli~\cite{eshost-cli}, also used at
Microsoft, for checking output discrepancy.\Mar{any difference in
  output is considered for reporting? are there any
  filtering?}

@@

\begin{wraptable}[10]{r}{0.22\textwidth}
%  \vspace{-2ex}      
  \centering
  \caption{\label{tab:test262}Percentage of passing tests on
    the Test262 conformance suite.}
  \vspace{-2ex}        
  \small
  \begin{tabular}{crr}
    \toprule
    engine & \% passing \\
    \midrule
    \jsc{} & 92\%\\
    \veight{} & 95\% \\
    \chakra{} & 75\% \\    
    \smonkey{} & 93\% \\
    \bottomrule
  \end{tabular}
  \normalsize
  \vspace{-4ex}
\end{wraptable}
For this experiment, we ran the suite once a day for seven consecutive
days and averaged the passing ratios. Table~\ref{tab:test262} shows
the average number of passing tests over this period. The variance of
results was negligible; for that reason, we omitted standard
deviations. We noticed that all four engines but \chakra{} used some
variant of the Test262 suite as part of their regression process, but
we used the same version in this
experiment~\cite{ecma262-conformance-suite}. Results show that there
are still many unsupported scenarios as can be observed from the
percentages in the table. The number of passing tests is high and
similar for \jsc{}, \veight{}, and \smonkey\ whereas \chakra\ performs
worse compared to the other engines. Note also that \chakra\ is both
the engine that has the lowest passing ratio in this test suite and
the one we were able to find more bugs (as per
Figure~\ref{fig:summary}). Although it is plausible to find
correlation between the passing ratios and reliability as measured by
the number of bugs found, we do not imply causality. It is important
to note that failures in this conformance test suite indicates missing
features as opposed to bugs.

@@

\Fix{Igor, please check these numbers! The status of most reports should
  have changed.}

@@

%% This suggests that the investiment in
%% testing paid off in this case.
\begin{table}[t]
  \small
  \centering
  \caption{\label{tab:cross-testing}Number of failures observed exchanging
    test suites.}
  \renewcommand*{\arraystretch}{0.9}
  \begin{tabular}{crrrr}
    \toprule
    test suite\textbackslash{}engine & \jsc{} & \veight{} & \smonkey{} & \chakra{}\\
    \midrule
    % \Fix{Versions (03.08)} & 234555 & 7.0.158 & 62.0b14 & 1.10.1 \\
    \Comment{
      Lembrar dos testes que os testes da propria engine falham:
      V8 0 
      JSC 2 
      Spidermonkey 58
    }
    \jsc{} & - & 9 & 10 & 58   \\
    \veight{} & 29 & - & 0 & 0  \\
%    \chakra{} & - & - & - & -  \\
    \smonkey{} & 218 & 107 & - & 279 \\
    Duktape & 0 & 4 & 4 & 1   \\
    \jerry{} & 23 & 25 & 22 & 23   \\
    JSI & 0 & 0 & 0 & 0   \\ 
   Tiny-js & 0 & 0 & 0 & 0  \\
    \midrule
   \textbf{total} & 270 & 145 & 36 & 361 \\
    \bottomrule 
  \end{tabular}
  \vspace{-3ex}
\end{table}

@@

\begin{wraptable}[12]{r}{0.23\textwidth}
  \vspace{-3ex}
  \small
  \centering
  \caption{\label{fig:falsepositives}\label{fig:truepositives}\label{fig:piecharts-transplantation}Distribution
    of False (FP) and True Positives (TP).}
  \setlength{\tabcolsep}{3pt}      
  \renewcommand*{\arraystretch}{0.9}
  \begin{tabular}{ccr}
    \toprule
    & source &  \#\\
    \midrule
    \multirow{5}{*}{FP} & Undefined Behavior & 252 \\
    & Timeout/OME & 15 \\
    & Not Implemented & 22 \\
    & Non-Standard Element & 119 \\
    & Other & 135 \\
    \midrule
    \multirow{2}{*}{TP} & Duplicate & 10 \\
    & Bug & 24 \\
    \bottomrule
  \end{tabular}
\end{wraptable}
Table~\ref{fig:piecharts-transplantation} shows the distribution of
False Positives (FPs) and True Positives (TPs). Considering false
positives, ``Undefined Behavior'' was the most predominant
source. Considering true positives, we found a reasonable number of
duplicate reports, but not high enough to justify attempting to
automate the detection of duplicates.

@@

\vspace{0.5ex}
\subsubsection{Results} Table~\ref{tab:summary-lo}
shows statistics of \lo\ warnings obtained when fuzzing inputs with
different tools. Column ``\# buckets'' shows the number of buckets
associated with ``\lo'' warnings, column ``\# files'' shows the total
number of files involved in these warnings, and column ``avg. \# files
per bucket'' shows the average number of files per bucket. Note that,
although the number of buckets and files differ substantially across
fuzzers, the difference of files per bucket was not very high.\Mar{not
sure how relevant is this par.}

\begin{table}[t]
  \small
  \centering
  \caption{\label{tab:summary-lo}Summary of \lo\ warning reports.}
  \begin{tabular}{crrr}
    \toprule
    & \# buckets & \# files & avg. \# files per bucket \\
    \midrule
    \radamsa{} & 82 & 319 & 3.8 \\
    \quickfuzz{} & 78 & 218 & 2.7 \\
    \bottomrule
  \end{tabular}
\end{table}

Table~\ref{tab:summary-hi} shows statistics of \hi\ warnings, which
are not clustered. The table breaks down \hi\ warning by the affected
engine, \ie, the engine manifesting distinct output among those
analyzed. Column ``+1'' shows the cases where more than one engine
disagree on the output. Note from the totals that the ordering of
engines is consistent with the one observed on
Table~\ref{tab:cross-testing}, with \chakra\ and \jsc\ in first and
second places, respectively, in number of warnings.

@@

\begin{table}[t]
  \small
  \centering
  \caption{\label{tab:false-positives}Distribution of False (FP)
    and True Positives (TP).\Mar{why numbers in this table does match with
      that on table ``Number of hi warnings...''?}}
  \begin{tabular}{ccrr}
    \toprule
    & & \radamsa\ & \quickfuzz\ \\
    \midrule
    \multirow{4}{*}{FP} & Undefined Behavior & 26 & 9 \\
    & Timeout/OME & 28 & 13 \\
    & * Invalid Input & 39 & 48 \\    
    & * Error Message Mismatch & 41 & 12 \\
    \midrule
    \multirow{2}{*}{TP} & Duplicate & 34 & 28\\
    & Bug & 12 & 6\\
    \bottomrule     
  \end{tabular}
\end{table}

@@

\subsection{Lessons Learned}
\label{sec:lessons}

We learned several lessons in this study:

\begin{itemize}
  \item The diversity-aware techniques we used are great
    at finding bugs in \js\ engines;
  \item The cost of inspecting failures/warnings is high. Although the
    inspection activity was not uninterrupted, it is safe to say that
    each warning required a substantial amount of time to analyze for
    potential false alarms;
  \item We have found more bugs with test transplantation than we have
    with cross-engine differential testing. In addition, test
    transplantation demanded less effort per bug;
  \item Even simple black-box fuzzers can create surprisingly
    interesting inputs;
  \item Mozilla's \smonkey\ appears to be the most reliable
    \js\ engine we used, with Google's \veight\ after it;
  \item Reporting real bugs is a great way to train (and encourage)
    students in software testing.
\end{itemize}

\Fix{revisor(es) reclamaram de pouca discussao. isto poderia ser usado
  em outros contexto? (resp: JS possui varias engine e muitos
  tests...) ver tambem ``key findings'' na introducao (precisa ficar
  consistente)}

%%%%%%%%%%%%%%%%%%%%%%%
%%% comentario de igor
%%%%%%%%%%%%%%%%%%%%%%%
%% sim, me refiro a testes que possuem algum comentário do que se trata aquele teste. Por exemplo os testes262, possuem comentários descrevendo o que o teste tem que fazer
%% os testes da suite do mozilla são mais dificeis de debuggar... não dificeis de grosso modo, mas custava um pouco para identificar o real problema

@@

>>>>

\emph{Key Findings.}~1) Not only multiple different implementations
can be leveraged in differential testing, but differences in test
suites can also be important. 2) Even for problems with fairly clear
specifications, as in \javascript{}, there is likely to be (a lot of)
variation between different implementations. 3) Differential testing
is feasible on real, complex, widely used pieces of
software. Section~\ref{sec:lessons} expands and elaborates key
findings and lessons learned.

@@

\vspace{1ex} To sum, the paper provides initial, yet strong evidence
that exploring diversity should be encouraged to find functional bugs
in JavaScript engines.

@@

\section{Mined JS Files}
\label{sec:seeds}
A good test set is critical for finding bugs with the techniques used
in this paper. For that reason, we looked for \js\ files from various
sources: 1) test files from the Test262~\cite{tc39-github} conformance
suite of the ECMA262 specification~\cite{ecmas262-spec}, 2) test files
from the test suite of the selected engines (see
Section~\ref{sec:methodology:engines}); these files are accessible
from the engine's official repositories, 3) test files from the suites
of public engines, and 4) test files mined from issue trackers of
these engines.

@@

This section describes the infrastructure we used for cross-engine
differential testing. Figure~\ref{fig:workflow} illustrate the
workflow of the approach. It takes on input a list of JS files and
generates warnings on output. Boxes in the figure denote
encapsulation, arrowed lines denote data flow, and the cycle icons
denote repetition--the leftmost icon indicates that each file in the
input list will be analyzed in separate whereas the rightmost icon
shows that a single file will be fuzzed multiple times.  The
bug-finding process works as follows. New inputs are obtained for a
given test input using some off-the-shelf input fuzzer;
Section~\ref{sec:objects:fuzzers} describes the fuzzers we selected.
The oracle checks whether or not the output produced for the fuzzed
file is consistent across all engines. In case the test passes in all
engines or fails in all engines (\ie{}, the output is consistent), the
infrastructure ignores the input. Otherwise, it considers the input as
potentially fault-revealing; hence, interesting for human
inspection. The infrastructure uses the open-source tool
eshost-cli~\cite{eshost-cli}, also used at Microsoft, for checking
output discrepancy.\Mar{any difference in output is considered for
  reporting? are there any filtering?}

@@

For this experiment, we ran the suite once a day for seven consecutive
days and averaged the passing ratios. Table~\ref{tab:test262} shows
\begin{wraptable}[10]{r}{0.22\textwidth}
  \vspace{-2ex}      
  \centering
  \caption{\label{tab:test262}Percentage of passing tests on
    the Test262 conformance suite.}
  \vspace{-2ex}        
  \small
  \begin{tabular}{crr}
    \toprule
    engine & \% passing \\
    \midrule
    \jsc{} & 92\%\\
    \veight{} & 95\% \\
    \chakra{} & 75\% \\    
    \smonkey{} & 93\% \\
    \bottomrule
  \end{tabular}
  \normalsize
  \vspace{-4ex}
\end{wraptable}
the average number of passing tests over this period. The variance of
results was negligible; for that reason, we omitted standard
deviations. We noticed that all four engines but \chakra{} used some
variant of the Test262 suite as part of their regression process, but
we used the same version in this
experiment~\cite{ecma262-conformance-suite}. Results show that there
are still many unsupported scenarios as can be observed from the
percentages in the table. The number of passing tests is high and
similar for \jsc{}, \veight{}, and \smonkey\ whereas \chakra\ performs
worse compared to the other engines. Note also that \chakra\ is both
the engine that has the lowest passing ratio in this test suite and
the one we were able to find more bugs (as per
Figure~\ref{fig:summary}). Although it is plausible to find
correlation between the passing ratios and reliability as measured by
the number of bugs found, we do not imply causality. It is important
to note that failures in this conformance test suite indicates missing
features as opposed to bugs.

@@

@@

%% This suggests that the investiment in
%% testing paid off in this case.
\begin{table}[t]
  \small
  \centering
  \caption{\label{tab:cross-testing}Number of failures with
    Test Transplantation.}
  \renewcommand*{\arraystretch}{0.9}
  \begin{tabular}{crrrr}
    \toprule
    test suite\textbackslash{}engine & \jsc{} & \veight{} & \smonkey{} & \chakra{}\\
    \midrule
    % \Fix{Versions (03.08)} & 234555 & 7.0.158 & 62.0b14 & 1.10.1 \\
    \Comment{
      Lembrar dos testes que os testes da propria engine falham:
      V8 0 
      JSC 2 
      Spidermonkey 58
    }
    \jsc{} & - & 9 & 10 & 58   \\
    \veight{} & 29 & - & 0 & 0  \\
%    \chakra{} & - & - & - & -  \\
    \smonkey{} & 218 & 107 & - & 279 \\
    Duktape & 0 & 4 & 4 & 1   \\
    \jerry{} & 23 & 25 & 22 & 23   \\
    JSI & 0 & 0 & 0 & 0   \\ 
   Tiny-js & 0 & 0 & 0 & 0  \\
    \midrule
   \textbf{total} & 270 & 145 & 36 & 361 \\
    \bottomrule 
  \end{tabular}
  \vspace{-3ex}
\end{table}

@@

\begin{wraptable}[12]{r}{0.23\textwidth}
  \vspace{-1ex}
  \small
  \centering
  \caption{\label{fig:falsepositives}\label{fig:truepositives}\label{fig:piecharts-transplantation}Distribution
    of False (FP) and True Positives (TP).}
  \setlength{\tabcolsep}{3pt}      
  \renewcommand*{\arraystretch}{0.9}
  \begin{tabular}{ccr}
    \toprule
    & source &  \#\\
    \midrule
    \multirow{5}{*}{FP} & Undefined Behavior & 252 \\
    & Timeout/OME & 15 \\
    & Not Implemented & 22 \\
    & Non-Standard Element & 119 \\
    & Other & 135 \\
    \midrule
    \multirow{2}{*}{TP} & Duplicate & 10 \\
    & Bug & 24 \\
    \bottomrule
  \end{tabular}
\end{wraptable}
Table~\ref{fig:piecharts-transplantation} shows the distribution of
False Positives (FPs) and True Positives (TPs). Considering false
positives, ``Undefined Behavior'' was the most predominant
source. Considering true positives, we found a reasonable number of
duplicate reports, but not high enough to justify attempting to
automate the detection of duplicates.

@@

%% Table~\ref{tab:summary-lo}
%% shows statistics of \lo\ warnings obtained when fuzzing inputs with
%% different tools. Column ``\# buckets'' shows the number of buckets
%% associated with ``\lo'' warnings, column ``\# files'' shows the total
%% number of files involved in these warnings, and column ``avg. \# files
%% per bucket'' shows the average number of files per bucket. Note that,
%% although the number of buckets and files differ substantially across
%% fuzzers, the difference of files per bucket was not very high.\Mar{not
%% sure how relevant is this par.}

%% \begin{table}[t]
%%   \small
%%   \centering
%%   \caption{\label{tab:summary-lo}Summary of \lo\ warning reports.}
%%   \begin{tabular}{crrr}
%%     \toprule
%%     & \# buckets & \# files & avg. \# files per bucket \\
%%     \midrule
%%     \radamsa{} & 82 & 319 & 3.8 \\
%%     \quickfuzz{} & 78 & 218 & 2.7 \\
%%     \bottomrule
%%   \end{tabular}
%% \end{table}

\vspace{0.5ex}
\subsubsection{Results}~Table~\ref{tab:summary-hi} shows statistics of \hi\ warnings. The table breaks down \hi\ warning by the affected
engine, \ie, the engine manifesting distinct output among those
analyzed. Column ``+1'' shows the cases where more than one engine
disagree on the output. Note from the totals that the ordering of
engines is consistent with the one observed on
Table~\ref{tab:cross-testing}, with \chakra\ and \jsc\ in first and
second places, respectively, in number of warnings.

@@

\begin{table}[t]
  \small
  \centering
  \caption{\label{tab:false-positives}Distribution of False (FP)
    and True Positives (TP).}
  \begin{tabular}{ccrr}
    \toprule
    & & \radamsa\ & \quickfuzz\ \\
    \midrule
    \multirow{4}{*}{FP} & Undefined Behavior & 26 & 9 \\
    & Timeout/OME & 28 & 13 \\
    & * Invalid Input & 39 & 48 \\    
    & * Error Message Mismatch & 41 & 12 \\
    \midrule
    \multirow{2}{*}{TP} & Duplicate & 34 & 28\\
    & Bug & 12 & 6\\
    \bottomrule     
  \end{tabular}
\end{table}

@@

\subsection{Key Findings and Lessons}
\label{sec:lessons}

The main findings of this study are as follows.

\begin{itemize}
  \item Both techniques investigated were successful in revealing
    bugs (several of them);
  %% \item We have found more bugs with test transplantation than we have
  %%   with cross-engine differential testing. In addition, test
  %%   transplantation demanded less effort per bug;
  \item Even simple black-box fuzzers can create surprisingly
    interesting inputs. We conjecture that there is room to find more
    bugs using other fuzzers;
  \item Mozilla's \smonkey\ appears to be the most reliable
    \js\ engine we used, with Google's \veight\ after it;
  \item Reporting real bugs is a great way to train (and encourage)
    students in software testing.
\end{itemize}

The main lessons we learned in this study are the following: 1)~Even
for problems with fairly clear specifications, as in \javascript{},
there is likely to be (a lot of) variation between different
implementations. 2)~Not only multiple different implementations can be
leveraged in differential testing, but differences in test suites can
also be important. 3)~Differential testing is feasible on real,
complex, widely used pieces of software. 4)~The cost of inspecting
failures/warnings is high. Although the inspection activity was not
uninterrupted, it is safe to say that each warning required a
substantial amount of time to analyze for potential false alarms. We
prefer to see this as an opportunity for future research in applying
learning techniques to prioritize warnings more likely to be faulty.
%based on static and dynamic features.

@@

\Fix{
although more research needs be done to automate
parts of the inspection process (\eg{}, reduction of false alarms),}
